---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Example: Modeling Full Expansion of Medicaid Under the ACA

The set of parameters encompassed by $\mathbf{c}_{t_0}$, $\mathbf{R(0)}$, and $\mathbf{R(1)}$ are sufficient statistics for modeling the impact of potential reforms to US health insurance markets. Critically, these parameters could be generated based on a microsimulation model, or they could be derived simply from reduced form estimates of the relevant quantities of interest. 

In this section, we demonstrate a reduced-form approach to modeling expansion of health insurance coverage. 

```{r sipp-setup, echo = FALSE, message = FALSE, warning = FALSE}
# https://www.nber.org/papers/w25568.pdf pp. 42
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)

scale_fill_Publication <- function(...){
  library(scales)
  discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
  
}

scale_colour_Publication <- function(...){
  library(scales)
  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
  
}

theme_tufte_revised <- function(base_size = 11, base_family = "Gill Sans", ticks = TRUE) {
  
  ret <- ggplot2::theme_bw(base_family = base_family, base_size = base_size) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = 'black'),
      axis.title.x = ggplot2::element_text(vjust = -0.3),
      axis.title.y = ggplot2::element_text(vjust = 0.8),
      legend.background = ggplot2::element_blank(),
      legend.key = ggplot2::element_blank(),
      legend.title = ggplot2::element_text(face="plain"),
      panel.background = ggplot2::element_blank(),
      panel.border = ggplot2::element_blank(),
      panel.grid = ggplot2::element_blank(),
      plot.background = ggplot2::element_blank(),
      strip.background = ggplot2::element_blank()
    )
  
  if (!ticks) {
    ret <- ret + ggplot2::theme(axis.ticks = ggplot2::element_blank())
  }
  
  ret
}
```

```{r sipp-run-functions}
# The primary code for this section can be found in "../../../box/sccs-r01/transitions-dd/analysis/multi-state-dd-modeling-primary-analysis.Rmd" 

source(here::here("../../../box/sccs-r01/transitions-dd/R/manifest.R"))
#devtools::session_info()
source(here("../../../box/sccs-r01/transitions-dd/R//prepare-multistate-data.R"))
source(here("../../../box/sccs-r01/transitions-dd/R//fit-multistate-model.R"))
source(here("../../../box/sccs-r01/transitions-dd/R//get-cumulative-hazard.R"))
source(here("../../../box/sccs-r01/transitions-dd/R//add-binary-indicators.R"))
source(here("../../../box/sccs-r01/transitions-dd/R//transitions-dd-model.R"))
source(here("../../../box/sccs-r01/transitions-dd/R//ri-permute.R"))

insurance_sipp_lut <- c(
  "cat_1" = "01_esi_own",
  "cat_2" = "02_esi_dep",
  "cat_3" = "03_priv_oth",
  "cat_4" = "04_public",
  "cat_5" = "05_uninsured"
)

insurance_lut <- 
  c("01_esi_own" = "Employer - Own Policy",
    "02_esi_dep" = "Employer - Dependent",
    "03_priv_oth" = "Private - Non-Employer",
    "04_public" = "Public",
    "05_uninsured" = "Uninsured")


conflict_prefer("filter","dplyr")
```

```{r}
months_to_model <- c(1,24) 
create_sipp_data = FALSE
if (create_sipp_data) source(here("../../../box/sccs-r01/transitions-dd/R/read-and-tidy-SIPP-data.R"))

df_sipp_full <- read_rds(here("../../../box/sccs-r01/transitions-dd/input/sipp/01_sipp-tidy_v2-0.rds"))

# Get the weight (first month weight)
df_w <- 
  df_sipp_full %>% 
  mutate(idnumber = id) %>% 
  group_by(idnumber) %>% 
  mutate(year = ifelse(swave ==1 , 2014, ifelse(swave==2,2015,2016))) %>% 
  mutate(month = ifelse(swave ==1 , monthcode, ifelse(swave==2,monthcode+12,monthcode+24))) %>% 
  filter(month==1) %>% 
  mutate(expansion_state = 
           as.integer(expansion_date == "2014-01-01" & 
                        (substantial_expansion_state == 1 | full_expansion_state == 1 |
                           mild_expansion_state==1))) %>% 
  mutate(exclude_state = as.integer(expansion_date != "2014-01-01")) %>% 
  # Expansion state varies in the underling data due to moves. Just keeping state at 
  # baseline. 
  select(idnumber,contains("state"), sex, race, weight = wpfinwgt) %>% 
  ungroup() 

df_sipp <- 
  df_sipp_full %>%   
  # Expansion state varies in the underling data due to moves. Just keeping state at 
  # baseline. 
  select(-contains("state")) %>% 
  filter(age < 63 & age > 18) %>% 
  mutate(year = ifelse(swave ==1 , 2014, ifelse(swave==2,2015,2016))) %>% 
  mutate(month = ifelse(swave ==1 , monthcode, ifelse(swave==2,monthcode+12,monthcode+24))) %>% 
  mutate(idnumber = id) %>% 
  mutate(insurance_type = ifelse(rhimth==2 & medcaidmth==0 & medcarmth==0 & ephi_t==0 & milmth==0 & privhimth==0,5,
                                 ifelse(medcaidmth==1 | medcarmth==1,4,
                                        ifelse((ephi_t==1 | milmth==1) & own_cov==1,1,
                                               ifelse((ephi_t==1 | milmth==1) & own_cov==0,2,
                                                      ifelse(privhimth==1,3,NA)))))) %>% 
  
  mutate(insurance_type=factor(insurance_type, labels = insurance_sipp_lut)) %>% 
  #mutate(insurance_type = as.numeric(factor(insurance_type))) %>% 
  select(idnumber,month,insurance_type)  %>% 
  group_by(idnumber) %>% 
  ungroup()  %>% 
  left_join(df_w,"idnumber") %>% 
  filter(exclude_state != 1 | is.na(exclude_state)) %>% 
  mutate(weight = 1) %>% 
  mutate(state = as.character(paste0(state))) %>% 
  group_by(idnumber) %>% 
  mutate(months_in_sample = n())

################################################################################
# Define estimation sample (Two-Period Balanced Panel Case)
################################################################################
df <-                                     # Data frame for estimation
  df_sipp %>%
  filter(month %in% months_to_model) %>% 
  filter(months_in_sample == months_to_model[2]) %>% 
  mutate(post_expansion = as.integer(month>12)) %>% 
  select(insurance_type,expansion_state,post_expansion,idnumber,month,state,weight) %>% 
  arrange(idnumber,month) %>% 
  filter(!is.na(insurance_type)) %>% 
  group_by(idnumber) %>% 
  filter(n()==2) %>% 
  group_by(idnumber) %>% 
  mutate(tmp = ifelse(row_number()==1, insurance_type,lag(insurance_type))) %>% 
  ungroup() %>% 
  mutate(baseline_insurance_type = factor(tmp, labels = insurance_sipp_lut)) %>% 
  select(-tmp)
################################################################################

df_sipp <- 
  df_sipp %>% 
  mutate(in_twoper_sample = as.integer(idnumber %in% unique(df$idnumber))) %>% 
  ungroup() 

dd_ests_over_time <- 
  2:24 %>% 
  map(~(df_sipp %>% 
          ungroup() %>% 
    filter(in_twoper_sample==1) %>% 
  count(month,insurance_type,expansion_state,wt = weight) %>% 
  group_by(month,expansion_state) %>% 
  mutate(pct = n/sum(n)) %>% 
  filter(month %in% c(1,.x)) %>% 
  select(-n) %>% 
  spread(month,pct) %>% 
  set_colnames(c("insurance_type","expansion_state","pre","post")) %>% 
  filter(!is.na(expansion_state)) %>% 
  mutate(change = post - pre) %>% 
  select(insurance_type,expansion_state,change) %>% 
  ungroup() %>% 
  mutate(expansion_state = factor(expansion_state,labels = c("nonexpansion","expansion"))) %>% 
  spread(expansion_state,change) %>% 
  mutate(dd = expansion-nonexpansion)
  )) %>% 
  bind_rows(.id = "comparison") %>% 
  mutate(foo = as.numeric(paste0(comparison))+1) 

```

```{r}

p <- dd_ests_over_time %>% 
  mutate(insurance2 = insurance_lut[insurance_type]) %>% 
  ggplot(aes(x = foo , y = dd * 100)) + geom_line(aes(colour = insurance2)) +
  ylab("Percentage point change relative to January 2013") + 
  xlab("Months Since January 2013") + 
  geom_vline(aes(xintercept = 13),lty=2) + 
  xlim(c(0,45)) + 
  annotate("text",x = 13.5,y = 4,label = "Primary Medicaid\nExpansion Date",hjust =0,cex = 3, family  = "Gill Sans") + 
  geom_hline(aes(yintercept = 0)) + theme_tufte_revised() +
  geom_hline(yintercept = c(-2,-1,1,2,3,4),colour = "lightgrey",lty=2) + 
  scale_y_continuous(breaks = -2:4, labels = paste0(format(round(-2:4),nsmall=1)))

p_final <- directlabels::direct.label(p,method=list("last.bumpup",fontfamily = "Gill Sans",cex = .8))  
```

```{r, fig.cap = "Difference-in-Diffrence Estimates on Change in Transition Probability, by Coverage Type and Time Since January 2013"}
p_final
```

```{r}

run_results = FALSE

if (run_results) {
  dd_lm <- 
    df %>% 
    fit_transitions_dd(outcome = insurance_type,
                       id = idnumber,
                       time = month,
                       ff = as.formula( ~ post_expansion + expansion_state + post_expansion * expansion_state),
                       ff_trans = as.formula( ~ post_expansion + expansion_state + baseline_insurance_type + 
                                                 post_expansion * baseline_insurance_type + post_expansion * expansion_state + 
                                                 expansion_state * baseline_insurance_type + 
                                                post_expansion * expansion_state * baseline_insurance_type),
                       term = "post_expansion:expansion_state"
    )
  
  dd_lm_ri <- df %>%
      ri_permute(m = 1000, 
                 treatment = expansion_state, 
                 cluster = state, 
                 idvar = idnumber,
                 outcome = insurance_type) %>% 
      do(fit_transitions_dd(df = . ,
                            outcome = insurance_type,
                         id = idnumber,
                         time = month,
                         ff = as.formula( ~ post_expansion + expansion_state + post_expansion * expansion_state),
                         ff_trans = as.formula( ~ post_expansion + expansion_state + baseline_insurance_type + 
                                                   post_expansion * baseline_insurance_type + post_expansion * expansion_state + 
                                                   expansion_state * baseline_insurance_type + 
                                                  post_expansion * expansion_state * baseline_insurance_type),
                         term = "post_expansion:expansion_state"
      ))
  
    list(dd_lm = dd_lm, 
         dd_lm_ri = dd_lm_ri)  %>% 
      write_rds("output/results/primary-results.rds")
}


results <- 
  read_rds(here("../../../box/sccs-r01/transitions-dd/output/results/primary-results.rds"))
dd_lm <- results$dd_lm
dd_lm_ri <- results$dd_lm_ri
```


```{r}

get_twoper_result <- function(df) {
  df %>% 
  select(term,starts_with("marginal")) %>% 
  na.omit() %>% 
  gather(estimate, value, -term) %>% 
  filter(term=="post_expansion:expansion_state") %>% 
  mutate(group = insurance_sipp_lut)  %>% 
  select(group,value)
}

get_quantile_ranking <- function(obs,inf) {
  inf %>% 
  group_by(replicate) %>% 
  bind_rows(
    obs %>% mutate(replicate=0) 
  ) %>% 
  group_by(group) %>% 
  mutate(quantile_rank = percent_rank(value)) %>% 
  filter(replicate == 0) %>% 
  select(group,quantile_rank)
}

get_pvalue <- function(obs,inf) {
  inf %>% 
  group_by(replicate) %>% 
  left_join(obs,"group") %>% 
  mutate(obs_lt_est = as.integer(abs(value.y) < abs(value.x))) %>% 
  group_by(group) %>% 
  summarise(p_value = mean(obs_lt_est)) 
}

get_transprob_result <- function(df) {
  df %>% 
  select(ex_ante,starts_with("DD")) %>% 
  na.omit() %>% 
  gather(estimate, value, -ex_ante) %>% 
  mutate(ex_post = estimate) %>% 
  mutate(ex_post = gsub("^DD_","",ex_post)) %>% 
  select(ex_ante,ex_post,value)
}

get_dd_marginal <- function(p1,p0,R1,DD,C,D) ((p1 %*% (C * R1) - p1*D) - (p0 %*% (C * (R1-DD)) - p0*D))

get_marginal_estimates <- 
  function(df, C = matrix(1,nrow = 5, ncol = 5), D = rep(1,5)) {
    
    # Regression Parameters (not going to output this now as decompositions won't match)
    regression_parameters <- 
      df %>% 
        select(term,starts_with("marginal")) %>% 
        filter(term == "post_expansion:expansion_state") %>% 
        gather(ex_ante,beta,-term) %>% 
        select(-term)
    
    p1 <- df %>% filter(!is.na(ex_ante)) %>% pull(p_Z1_T0) #%>% as.matrix()
    p0 <- df %>% filter(!is.na(ex_ante)) %>% pull(p_Z0_T0) #%>% as.matrix()
    tmp <- df %>% filter(!is.na(ex_ante)) %>% select(ex_ante,starts_with("DD")) 
    DD <- tmp[,-1] %>% as.matrix()
    rownames(DD) <- tmp$ex_ante
    tmp2 <- df %>% filter(!is.na(ex_ante)) %>% select(ex_ante,starts_with("R1"))
    R1 <- tmp2[,-1] %>% as.matrix()
    rownames(R1) <- tmp2$ex_ante

    get_dd_marginal(p1 = p1, p0 = p0, R1= R1, DD = DD, D = D, C = C ) %>% 
      data.frame() %>% 
      gather(estimate,value) %>% 
      mutate(group = gsub("R1_","",estimate)) %>% 
      select(-estimate)
  }

get_decomp_est <- function(df,df_ri,C,D) {

  # Pooled (Transition) Model Marginal Result
  pooled_marginal_result_obs <- 
    df %>% 
    do(get_marginal_estimates(df = .,C = C, D = D)) 
  pooled_marginal_result_inf <- 
    df_ri %>% 
    group_by(replicate) %>% 
    do(get_marginal_estimates(df = .,C = C, D = D)) 
  pooled_marginal_result <- 
    pooled_marginal_result_obs %>% 
    left_join(get_quantile_ranking(obs = pooled_marginal_result_obs, inf = pooled_marginal_result_inf),"group") %>% 
    left_join(get_pvalue(obs = pooled_marginal_result_obs, inf = pooled_marginal_result_inf),"group")
  
  pooled_marginal_result_formatted <- 
    pooled_marginal_result %>% 
    select(group, a_pooledl = value, b_p_value = p_value) %>% 
    gather(row,value,-group) %>% 
    arrange(group) %>% 
    mutate(value = ifelse(grepl("^a_",row),format(round(value,4),nsmall=3),
                          ifelse(value>0.01,paste0("(p=",format(round(value,2),nsmall=1),")"),
                                 paste0("(p< 0.01)")))) %>% 
    mutate(group = ifelse(grepl("^a_",row),insurance_lut[group],"")) %>% 
    select(group,pooled=value)
  
  out <- 
    list(marginal_result = pooled_marginal_result, 
         marginal_result_formatted = pooled_marginal_result_formatted)
  
  return(out)
  
}


```

```{r}
## Ex Ante Shares
ex_ante_shares <- 
  dd_lm %>% 
  filter(!is.na(ex_ante)) %>% 
  select(ex_ante,starts_with("p_Z1_T0"), starts_with("p_Z0_T0")) %>% 
  na.omit() %>% 
  tbl_df() %>% 
  mutate_at(vars(2:3),function(x) format(round(x,3),nsmall=3)) %>% 
  mutate(test1 = "",test2="") %>% 
  gather(key,value,-ex_ante,-p_Z1_T0,-p_Z0_T0) %>% 
  arrange(ex_ante) %>% 
  group_by(ex_ante) %>% 
  mutate(n = row_number()) %>% 
  mutate(treated = ifelse(n==1,p_Z1_T0,NA),
         control = ifelse(n==1,p_Z0_T0,NA),
         group = ifelse(n==1,insurance_lut[ex_ante],NA)) %>% 
  ungroup() %>% 
  select(group,treated,control)
  
# Marginal Model Result

marginal_result_obs <- 
  dd_lm %>% 
  get_twoper_result()
marginal_result_inf <- 
  dd_lm_ri %>% 
  group_by(replicate) %>% 
  do(get_twoper_result(df = .)) 
marginal_result <- 
  marginal_result_obs %>% 
  left_join(get_quantile_ranking(obs = marginal_result_obs, inf = marginal_result_inf),"group") %>% 
  left_join(get_pvalue(obs = marginal_result_obs, inf = marginal_result_inf),"group")

marginal_result_formatted <- 
  marginal_result %>% 
  select(group, a_marginal = value, b_p_value = p_value) %>% 
  gather(row,value,-group) %>% 
  arrange(group) %>% 
  mutate(value = ifelse(grepl("^a_",row),format(round(value,4),nsmall=3),
                        ifelse(value>0.01,paste0("(p=",format(round(value,2),nsmall=1),")"),
                               paste0("(p< 0.01)")))) %>% 
  mutate(group = ifelse(grepl("^a_",row),insurance_lut[group],"")) %>% 
  select(group,marginal=value)


# Pooled (Transition) Model Marginal Result
pooled_marginal_result_obs <- 
  dd_lm %>% 
  do(get_marginal_estimates(df = .)) 
pooled_marginal_result_inf <- 
  dd_lm_ri %>% 
  group_by(replicate) %>% 
  do(get_marginal_estimates(df = .)) 
pooled_marginal_result <- 
  pooled_marginal_result_obs %>% 
  left_join(get_quantile_ranking(obs = pooled_marginal_result_obs, inf = pooled_marginal_result_inf),"group") %>% 
  left_join(get_pvalue(obs = pooled_marginal_result_obs, inf = pooled_marginal_result_inf),"group")
 
pooled_marginal_result_formatted <- 
  pooled_marginal_result %>% 
  select(group, a_pooledl = value, b_p_value = p_value) %>% 
  gather(row,value,-group) %>% 
  arrange(group) %>% 
  mutate(value = ifelse(grepl("^a_",row),format(round(value,4),nsmall=3),
                        ifelse(value>0.01,paste0("(p=",format(round(value,2),nsmall=1),")"),
                               paste0("(p< 0.01)")))) %>% 
  mutate(group = ifelse(grepl("^a_",row),insurance_lut[group],"")) %>% 
  select(group,pooled=value)

# Transition Probability Matrix Result
transprob_result_obs <- 
  dd_lm %>% 
  get_transprob_result() %>% 
  unite(group,ex_ante,ex_post,sep = "_TO_")

transprob_result_inf <- 
  dd_lm_ri %>% 
  group_by(replicate) %>% 
  do(get_transprob_result(df=.) %>% 
  unite(group,ex_ante,ex_post,sep = "_TO_")) 
  
transprob_result <- 
  transprob_result_obs %>% 
  left_join(get_quantile_ranking(obs = transprob_result_obs, inf = transprob_result_inf),"group") %>% 
  left_join(get_pvalue(obs = transprob_result_obs, inf = transprob_result_inf),"group")

transprob_result_formatted <-
 transprob_result %>% 
  select(group, a_transprobl = value, b_p_value = p_value) %>% 
  gather(row,value,-group) %>% 
  arrange(group) %>% 
  mutate(value = ifelse(grepl("^a_",row),format(round(value,4),nsmall=3),
                        ifelse(value>0.01,paste0("(p=",format(round(value,2),nsmall=1),")"),
                               paste0("(p< 0.01)")))) %>% 
   separate(group,into = c("ex_ante","ex_post"),sep = "_TO_") %>% 
   spread(ex_post,value) %>% 
  mutate(ex_ante = ifelse(grepl("^a_",row),insurance_lut[ex_ante],"")) %>% 
  select(-row) %>% 
   set_colnames(c("group",insurance_lut))


```


```{r}
tmp <- ex_ante_shares %>% 
  #cbind(marginal_result_formatted[,-1])  %>% 
  #cbind(pooled_marginal_result_formatted[,-1]) %>% 
  cbind(transprob_result_formatted[,-1]) 

tmp %>% kable()
```

- Get the transition probability matrix for the untreated states. 
- Get the DD transition probability matix.
- Add together to get the counterfactual. 
- Multiply by baseline occupancy vector.
- Assign costs to transition.
- Aggregate costs and population figures.
